---
title: "Stats Project"
output: html_document
date: "2024-10-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readr)
library(ggplot2)
library(dplyr)

# Adding postseason scores
cbb5 <- cbb %>% mutate(postseason_score = case_when(
  POSTSEASON == "R68" ~ 1,
  POSTSEASON == "R64" ~ 2,
  POSTSEASON == "R32" ~ 3,
  POSTSEASON == "S16" ~ 4,
  POSTSEASON == "E8" ~ 5,
  POSTSEASON == "F4" ~ 6,
  POSTSEASON == "2ND" ~ 7,
  POSTSEASON == "Champions" ~ 8,
  .default = 0
))

# Summing postseason scores for each team
cbb2 <- cbb5 %>% group_by(TEAM) %>% summarise(pst = sum(postseason_score, na.rm = TRUE))

# Merging with the adjusted dataset
cbb3 <- cbb_adj %>% left_join(cbb2, by = "TEAM")

# Preview the result
head(cbb3)

```
```{r}

#Filters out teams that have not been in the same conference for 10 years
cbbCONF <- cbb %>% 
  group_by(TEAM, CONF) %>%
  count() %>%  
  filter(n==10)

head(cbb)
  
```


```{r}
#Joins the filtered out conferences with the average stats and post season score. 
cbbjoined <- cbbCONF %>% 
  inner_join(cbb3)

head(cbbjoined)

```
```{r}
cbb_new<-cbbjoined%>%
  group_by(CONF)%>%
summarize(pst=mean(pst))%>%
  inner_join(cbbjoined)
```

```{r}


```

```{r}


all_offense_mod <- (lm(pst ~ EFG_O + TOR + ORB + FTR + `2P_O` + `3P_O`, cbbjoined))
all_defense_mod <- (lm(pst ~ EFG_D + TORD + DRB + FTRD + `2P_D` + `3P_D`, cbbjoined))
all_variable_mod<-(lm(pst ~ EFG_O+EFG_D+TOR+TORD+ORB+DRB+FTR+FTRD+`2P_O`+`2P_D`+`3P_O`+`3P_D`,cbbjoined))



model1<-lm(formula = pst ~ TOR, data = cbbjoined)
model2<-lm(formula = pst ~ TORD+ADJDE+EFG_D+DRB+FTRD, data = cbbjoined)
model3<-lm(formula=pst~TOR+CONF,data=cbbjoined)
model4<-lm(formula=pst~TORD+CONF,data=cbbjoined)
summary(model1)
summary(model2)
summary(model3)
summary(model4)
summary(all_offense_mod)
summary(all_defense_mod)
summary(all_variable_mod)

ggplot(all_variable_mod,aes(x=EFG_O,y=pst))+
  geom_point()+
  geom_smooth(method="lm",formula=y~x,se=FALSE)

ggplot(cbbjoined, aes(x = TOR, y = pst)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ x, se = FALSE) +
  labs(title = "Scatterplot of pst vs TOR",
       x = "Turnovers per game(TOR)",
       y = "Performance Score (pst)") +
  theme_minimal()

ggplot(cbbjoined, aes(x = TORD, y = pst)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ x, se = FALSE) +
  labs(title = "Scatterplot of pst vs TORD",
       x = "Turnovers forced per game(TORD)",
       y = "Performance Score (pst)") +
  theme_minimal()


ggplot(cbbjoined, aes(x = TOR, y = pst,color=CONF)) + 
  geom_point() + 
  geom_smooth(method = "lm", formula = y ~ x, se = FALSE) + 
  labs(
    title = "Scatterplot of pst vs TOR by Conference",
    x = "Turnovers per game (TOR)",
    y = "Wins(W)"
  ) +
facet_wrap(~CONF)


ggplot(cbbjoined, aes(x = TORD, y = pst,color=CONF)) + 
  geom_point() + 
  geom_smooth(method = "lm", formula = y ~ x, se = FALSE) + 
  labs(
    title = "Scatterplot of pst vs TORD by Conference",
    x = "Turnovers forced per game (TORD)",
    y = "Wins(W)"
  ) +
facet_wrap(~CONF)








```

```{r}
#Zachs
#Offense Only Model

lm_mod <- lm(pst ~ EFG_O, cbbjoined)
summary(lm_mod)

cbbjoined %>% 
  ggplot(aes(x=EFG_O, y=pst))+
  geom_point()+
  geom_smooth(se = F, method= "lm")+
  labs(x = "Effective Field Goal % Shot", y= "Postseason Total Score")
```

```{r}
#Defense Only Model
defense_mod <- lm(pst ~ EFG_D, cbbjoined)
summary(defense_mod)

cbbjoined %>% 
  ggplot(aes(x=EFG_D, y=pst))+
  geom_point()+
  geom_smooth(se = F, method= "lm")
```

```{r}
#Check Conference additive term
conference_modOff <- lm(pst ~ EFG_O + CONF, cbbjoined)
summary(conference_modOff)

conference_modDef <- lm(pst ~ EFG_D + CONF, cbbjoined)
summary(conference_modDef)

```
```{r}
#Conference facet wrapped graph

cbbjoined %>% 
  ggplot(aes(x = EFG_O, y = pst, color = CONF)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(~ CONF)+
  theme_minimal()+
  labs(x = "Effective Field Goal %", y= "Post Season Total")

cbbjoined %>% 
  ggplot(aes(x = EFG_D, y = pst, color = CONF)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(~ CONF)+
  theme_minimal()+
  labs(x = "Effective Field Goal % Allowed ", y= "Post Season Total")
```


```
```{r}

ggplot(cbbjoined,aes(x=`3P_O`,y=pst))+
  geom_point()+
  geom_smooth(method = "lm", se=FALSE)+
  labs(x = '3-Point Percentage',y = 'Postseason Score',title='3-Point Percentage vs Postseason Score')
ggplot(cbbjoined,aes(x=`3P_D`,y=pst))+
  geom_point()+
  geom_smooth(method = "lm", se=FALSE)+
  labs(x = '3-Point Percentage Allowed',y = 'Postseason Score',title='3-Point Percentage Allowed vs Postseason Score')

ThreePOModel<-lm(pst~`3P_O`,data=cbbjoined)
summary(ThreePOModel)
ThreePDModel<-lm(pst~`3P_D`,data=cbbjoined)
summary(ThreePDModel)
```

```{r}
#Nicks code
linear_model <- lm(pst ~ FTR +FTRD, cbbjoined)
coef(linear_model)

linear_model_2 <- lm(pst ~ W, cbbjoined)
coef(linear_model_2)

ggplot(cbbjoined, aes(x = FTR, y = pst)) +
  geom_point(alpha = 0.6, color = "blue") +
  geom_smooth(method = "lm", formula = y ~ x, se = FALSE, color = "red") +
  labs(title = "Scatterplot of pst vs FTR",
       x = "Free Throw Rate (FTR)",
       y = "Performance Score (pst)") +
  theme_minimal()

selected_ticks <- seq(min(cbbjoined$W), max(cbbjoined$W), length.out = 6)

ggplot(cbbjoined, aes(x = as.factor(W), y = pst)) +
  geom_boxplot(fill = "lightblue", color = "darkblue") +
  labs(title = "Boxplot of pst by Wins (W)",
       x = "Wins (W)",
       y = "Performance Score (pst)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 0, hjust = 1)) +
  scale_x_discrete(breaks = as.character(round(selected_ticks)))


```
```{r}

```

```
```{r}
cbbjoined <- cbbjoined %>%
  mutate(PowerSix = ifelse(CONF %in% c("B10", "B12", "BE", "SEC", "P12", "ACC"), "Power Six", "Other"))

cbbjoined %>%
  ggplot(aes(x = EFG_D, y = pst, color = PowerSix)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(~ PowerSix) +
  theme_minimal() +
  labs(
    x = "Effective Field Goal % Allowed",
    y = "Post Season Total",
    color = "Conference Type",
    title = "Comparison of Power Six vs Other Conferences"
  )


cbbjoined <- cbbjoined %>% ungroup() %>% 
  mutate(FG_zscore = scale(EFG_O),
         TOR_zscore = scale(TOR), 
         ORB_zscore = scale(ORB),
         FTR_zscore = scale(FTR),
         twoPFG_zscore = scale(`2P_O`),
         threePFG_zscore = scale(`3P_O`),
         Total_zscoreO = rowSums(across(c(FG_zscore, TOR_zscore, ORB_zscore, FTR_zscore, twoPFG_zscore, threePFG_zscore)))
)

cbbjoined <- cbbjoined %>% ungroup() %>% 
  mutate(FGD_zscore = scale(EFG_D),
         TORD_zscore = scale(TORD), 
         DRB_zscore = scale(DRB),
         FTRD_zscore = scale(FTRD),
         twoPFGD_zscore = scale(`2P_D`),
         threePFGD_zscore = scale(`3P_D`),
)
```

